The role of lithic bipolar technology in Western Iberia’s Upper Paleolithic: the case of Vale Boi (southern Portugal)

### 2018-06-08 analysis/paper/paper.Rmd
unused functions deleted

source('~/R_DATA/compendiums/ScaledPiecesVB/R/scaled_pieces_functions.R')

@@

all_data <- read_data()
techno_data <- all_data$techno_data
morpho_data <- all_data$morpho_data

@@

assemb <- cross_tb(techno_data, "RawMaterial", var_list)

@@

Other characteristics of the assemblage are the low frequency of retouch (n = `r techno_data %>% filter(RetouchedBlank == "Yes") %>% summarise(n())`), of fire alterations (n = `r techno_data %>% filter(Fire != "No_Traces") %>% summarise(n())`), and of the presence of original striking platforms (n = `r techno_data %>% filter(ButtType != "Absent") %>% summarise(n())`). Still, when present, retouch is located in the lateral part of the artifacts, similar to what would define a side-scraper. In a very small number of cases, when striking platforms are present, these are mostly flat. The absence of the original blank platforms is to be expected in this type of artifact, mostly due to the functional use of the pieces, and consequent removal of the platform, rather than an actual choice. The large absence of platforms may, also, be the result of the use of the technological axis as main functional axis. In fact, when identifiable, technological and typological axes coincide in `r inline_perc(techno_data, AxisCoincident, "Yes")`% of the cases. The longitudinal sections show a large variability of shapes, independent of raw materials or chronologies. Similarly, the dorsal pattern of previous removals was rarely identified, although this, like with the case of platform absence, might occur due to the functional stigmas and be dependent on the intensity of use for each artifact.

@@

# Get cross-tables by chronology using cross_tb() function
Gravettian <- cross_tb(filter$Gravettian, "RawMaterial", var_list) 
Gravettian <- Gravettian %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Total = Overall)

Solutrean <- cross_tb(filter$Solutrean, "RawMaterial", var_list)
Solutrean <- Solutrean %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Chalcedony, Total = Overall)

`Proto-Solutrean` <- cross_tb(filter$`Proto-Solutrean`, "RawMaterial", var_list)
`Proto-Solutrean` <- `Proto-Solutrean` %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Chalcedony, Total = Overall)

Magdalenian <- cross_tb(filter$Magdalenian, "RawMaterial", var_list)
Magdalenian <- Magdalenian %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Total = Overall)

@@

area <- mean_plot(techno_data, "Chronology", "Area", "RawMaterial")
thickness <- mean_plot(techno_data, "Chronology", "Thickness", "RawMaterial")

@@

area_by_platforms <- box_plot(techno_data, "DamagedPlatforms", "Area", "Number of Damaged Platforms", "Area (Length * Width)")
thickness_by_platforms <- box_plot(techno_data, "DamagedPlatforms", "Thickness", "Number of Damaged Platforms", "Thickness (mm)")

@@

#Select and arrange morpho-functional variables for each pair of opposed platforms
morpho_data_opposed_platforms <- opposed_plat_morpho()

@@

plot_morpho_var(morpho_data_opposed_platforms)

@@

To identify possible patterns of association among the qualitative variables used in our analysis \@ref(fig:morphoattributes), a Multiple Correspondence Analysis was performed. Results are presented in Figures \@ref(fig:mca). In this analysis raw material was used as a qualitative supplementary variable.

@@

mca.rawmaterial()

>>>>

source('~/R_DATA/compendiums/ScaledPiecesVB/R/functions.R')

@@

all_data <- read.data()
techno_data <- all_data$techno_data
morpho_data <- all_data$morpho_data

@@

assemb <- cross.tb(techno_data, "RawMaterial", var_list)

@@

Other characteristics of the assemblage are the low frequency of retouch (n = `r techno_data %>% filter(RetouchedBlank == "Yes") %>% summarise(n())`), of fire alterations (n = `r techno_data %>% filter(Fire != "No_Traces") %>% summarise(n())`), and of the presence of original striking platforms (n = `r techno_data %>% filter(ButtType != "Absent") %>% summarise(n())`). Still, when present, retouch is located in the lateral part of the artifacts, similar to what would define a side-scraper. In a very small number of cases, when striking platforms are present, these are mostly flat. The absence of the original blank platforms is to be expected in this type of artifact, mostly due to the functional use of the pieces, and consequent removal of the platform, rather than an actual choice. The large absence of platforms may, also, be the result of the use of the technological axis as main functional axis. In fact, when identifiable, technological and typological axes coincide in `r inline.perc(techno_data, AxisCoincident, "Yes")`% of the cases. The longitudinal sections show a large variability of shapes, independent of raw materials or chronologies. Similarly, the dorsal pattern of previous removals was rarely identified, although this, like with the case of platform absence, might occur due to the functional stigmas and be dependent on the intensity of use for each artifact.

@@

# Get cross-tables by chronology using cross_tb() function
Gravettian <- cross.tb(filter$Gravettian, "RawMaterial", var_list) 
Gravettian <- Gravettian %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Total = Overall)

Solutrean <- cross.tb(filter$Solutrean, "RawMaterial", var_list)
Solutrean <- Solutrean %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Chalcedony, Total = Overall)

`Proto-Solutrean` <- cross.tb(filter$`Proto-Solutrean`, "RawMaterial", var_list)
`Proto-Solutrean` <- `Proto-Solutrean` %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Chalcedony, Total = Overall)

Magdalenian <- cross.tb(filter$Magdalenian, "RawMaterial", var_list)
Magdalenian <- Magdalenian %>% 
  as.tibble() %>% 
  select(" " = Variable, Chert, Quartz, Total = Overall)

@@

area <- mean.plot(techno_data, "Chronology", "Area", "RawMaterial")
thickness <- mean.plot(techno_data, "Chronology", "Thickness", "RawMaterial")

@@

area_by_platforms <- box.plot(techno_data, "DamagedPlatforms", "Area", "Number of Damaged Platforms", "Area (Length * Width)")
thickness_by_platforms <- box.plot(techno_data, "DamagedPlatforms", "Thickness", "Number of Damaged Platforms", "Thickness (mm)")

@@

#Select and arrange morpho-functional variables for each pair of opposed platforms
morpho_data_opposed_platforms <- opposed.plat.morpho()

@@

plot.morpho.var(morpho_data_opposed_platforms)

@@

To identify possible patterns of association among the qualitative variables used in our analysis \@ref(fig:morphoattributes), a Multiple Correspondence Analysis (MCA) was performed. Results are presented in Figures \@ref(fig:mca). In this analysis, raw material was used as a qualitative supplementary variable.

@@

morpho_data_mca <- mca.scaled.pieces()
